---
title: "R Notebook"
output: "HTML"
---

load libraries

```{r}
library(dplyr)
library(caret)
```

load in the data sets

```{r}
winLossData <- read.csv("compiled-features/allteams-allseason-allcoaches-winloss-1985-2017.csv", header = TRUE)
seasonAveragesData <- read.csv("compiled-features/allteams-allseason-averages.csv", header = TRUE)
seasonSdData <- read.csv("compiled-features/allteams-allseason-SD.csv", header = TRUE)
gamePossessionData <- read.csv("compiled-features/allteams-allseason-AveragePossessionsperGame.csv")
```


Create data base of features

```{r}
#prefix the column names of the averages and SD data sets
colnames(seasonAveragesData)[c(3:16)] <- paste("Avg", colnames(seasonAveragesData[,c(3:16)]), sep = "_")
colnames(seasonSdData)[c(3:16)] <- paste("Sd", colnames(seasonSdData[,c(3:16)]), sep = "_")

#begin creating the training set with the averages data filtered to 2010:2017
seasonData <- seasonAveragesData %>%
  filter(Season %in% c(2010:2017))

#filter the Sd data set to the same time period
seasonSdData <- seasonSdData %>%
  filter(Season %in% c(2010:2017))

#select columns to keep from win loss data set
winLossSelectionVector <- c(1:2,11:13,15)

#subset data set to only include columns above
winLossData <- winLossData[,winLossSelectionVector]

#filter win loss data set to same time period
winLossData <- winLossData %>%
  filter(Season %in% c(2010:2017))

#join all of the data
seasonData <- left_join(seasonData,seasonSdData)
seasonData <- left_join(seasonData,gamePossessionData)
seasonData <- left_join(seasonData,winLossData)

head(seasonData)
```

Create training set from season data

```{r}
seasonResults <- read.csv("data/RegularSeasonDetailedResults.csv", header = TRUE)
head(seasonResults)
```

```{r}
#subset 2014 data into new df
trainingSet <- seasonResults %>%
  filter(Season %in% c(2010:2017)) %>%
  select(Season, WTeamID, LTeamID)

#change the column names so that they are neutral (not indicative of winner or loser)
names(trainingSet) <- gsub(x = names(trainingSet), pattern = "WTeam", replacement = "Team1")
names(trainingSet) <- gsub(x = names(trainingSet), pattern = "LTeam", replacement = "Team2")
names(trainingSet) <- gsub(x = names(trainingSet), pattern = "W", replacement = "Team1")
names(trainingSet) <- gsub(x = names(trainingSet), pattern = "L", replacement = "Team2")



head(trainingSet)

```

```{r}
#pull data for team 1
trainingSet <- left_join(trainingSet,seasonData, by = c("Team1ID" = "TeamID", "Season" = "Season"))

# rename variables for team1
colnames(trainingSet)[c(4:38)] <- paste("Team1", colnames(trainingSet[,c(4:38)]), sep = "_")

#pull data for team 1
trainingSet <- left_join(trainingSet,seasonData, by = c("Team2ID" = "TeamID", "Season" = "Season"))

# rename variables for team1
colnames(trainingSet)[c(39:73)] <- paste("Team2", colnames(trainingSet[,c(39:73)]), sep = "_")

#create a results variable
trainingSet <- trainingSet  %>% 
  mutate(Result = 1)
```


```{r}
#duplicate data
trainingSetReverse <- trainingSet

#reverse the result
trainingSetReverse <- trainingSetReverse %>%
     mutate(Result = replace(Result, Result == 1, 0)) %>%
     as.data.frame()

#rename Team 1 to Team 2 and vice versa
names(trainingSetReverse) <- gsub(x = names(trainingSetReverse), pattern = "Team1", replacement = "temp1")
names(trainingSetReverse) <- gsub(x = names(trainingSetReverse), pattern = "Team2", replacement = "temp2")
names(trainingSetReverse) <- gsub(x = names(trainingSetReverse), pattern = "temp1", replacement = "Team2")
names(trainingSetReverse) <- gsub(x = names(trainingSetReverse), pattern = "temp2", replacement = "Team1")

#reorder columns manually (TODO: this could probably be done better)
trainingSetReverse <- trainingSetReverse[,c(1,3,2,39:73,4:38,74)]
head(trainingSetReverse)

#bind datasets together
trainingSet <- rbind(trainingSet,trainingSetReverse)

trainingSet <- trainingSet[complete.cases(trainingSet),]

#write.csv(trainingSet, file = "trainingSet.csv", row.names = FALSE)

```



```{r}
#it appears that all of the columns have values between 0 and 1, but we apply a simple code to feature scale for good practice
for (i in c(4:73)){
 trainingSet[,i] <- (trainingSet[,i] - mean(trainingSet[,i]))/sd(trainingSet[,i])
 trainingSet[,i] <- (trainingSet[,i] - mean(trainingSet[,i]))/sd(trainingSet[,i])
}

head(trainingSet)
```

```{r}
trainingSet$Result <- as.integer(trainingSet$Result)
summary(trainingSet$Result)

numSamples <- nrow(trainingSet)
sampleRate <- 0.015
trainingRows <- sample(1:numSamples, sampleRate*numSamples, replace = FALSE)
trainingSet <- trainingSet[trainingRows,]


```

```{r}
#trainingSet <- trainingSet[,c(1:5,12:13,32:35,38:40,47:48,67,70,73,74)]


trainingSet <- trainingSet[,c(4:74)]
head(trainingSet)
```


Win Per
Win per not Home
Avg Score
Total Possession Time
AVG FGM
AVG AST
AVG DR
AVG TO
AVG FTM
AVG BLK
SD AST

Team1_WinPer + Team2_WinPer + Team1_WinPerNotHome + Team2_WinPerNotHome + Team1_Avg_Score + Team2_Avg_Score + Team1_TotalPossessionTime + Team2_TotalPossessionTime + Team1_Avg_FGM + Team2_Avg_FGM + Team1_Avg_Ast + Team2_Avg_Ast + Team1_Avg_DR + Team2_Avg_DR

```{r}
myform <- as.formula(paste0('Result ~ ', 
                             paste(names(trainingSet[!names(trainingSet) %in% 'Result']),
                     collapse = ' + ')))
```


##Build NN
```{r}
library(neuralnet)

nnModel <- neuralnet(myform, data = trainingSet, hidden = c(9,5,3), linear.output = FALSE)

```

```{r}
plot(nnModel)
```

Create TESTING set from season data

```{r}
tourneyResults <- read.csv("data/NCAATourneyDetailedResults.csv", header = TRUE)
head(tourneyResults)
```

```{r}
#subset 2014 data into new df
testingSet <- tourneyResults %>%
  filter(Season %in% c(2010:2017)) %>%
  select(Season, WTeamID, LTeamID)

#change the column names so that they are neutral (not indicative of winner or loser)
names(testingSet) <- gsub(x = names(testingSet), pattern = "WTeam", replacement = "Team1")
names(testingSet) <- gsub(x = names(testingSet), pattern = "LTeam", replacement = "Team2")
names(testingSet) <- gsub(x = names(testingSet), pattern = "W", replacement = "Team1")
names(testingSet) <- gsub(x = names(testingSet), pattern = "L", replacement = "Team2")



```




```{r}
#pull data for team 1
testingSet <- left_join(testingSet,seasonData, by = c("Team1ID" = "TeamID", "Season" = "Season"))

# rename variables for team1
colnames(testingSet)[c(4:38)] <- paste("Team1", colnames(testingSet[,c(4:38)]), sep = "_")

#pull data for team 1
testingSet <- left_join(testingSet,seasonData, by = c("Team2ID" = "TeamID", "Season" = "Season"))

# rename variables for team1
colnames(testingSet)[c(39:73)] <- paste("Team2", colnames(testingSet[,c(39:73)]), sep = "_")

#create a results variable
testingSet <- testingSet  %>% 
  mutate(Result = NA)

```

```{r}
#keep only columns in the model
#testingSet <- testingSet[,c(1:5,12:13,32:35,38:40,47:48,67,70,73,74)]


#testingSet <- testingSet[,c(1:3,35,70,38,73,4,39,32,67,5,40,13,48,12,47,74)]
testingSet <- testingSet[,c(4:74)]

head(testingSet)



```


```{r}
#drop certain rows that are not needed

#it appears that all of the columns have values between 0 and 1, but we apply a simple code to feature scale for good practice
for (i in c(1:70)){
 testingSet[,i] <- (testingSet[,i] - mean(testingSet[,i]))/sd(testingSet[,i])
 testingSet[,i] <- (testingSet[,i] - mean(testingSet[,i]))/sd(testingSet[,i])
}

head(testingSet)

numSamples <- nrow(testingSet)
sampleRate <- 0.8
testingRows <- sample(1:numSamples, sampleRate*numSamples, replace = FALSE)
testingSet <- testingSet[testingRows,]

head(testingSet)
```



Test NN model

```{r}

# Perform prdictions for the testing set
predictedLabels <-compute(nnModel, testingSet[,1:70])
#The result is stored in predictedLabels$net.result
#Note that we need to round the predictions to get 0 or 1
#predictedLabels<-round(predictedLabels$net.result)
#predictedLabels


#predictedLabels <- round(predictions)
#predictedLabels

predictedLabels <- as.data.frame(predictedLabels)
```

```{r}

```


```{r}
library(MLmetrics)

LogLoss(y_pred = predictedLabels[,92], y_true = 1)

predictedLabels<-round(predictedLabels$net.result)

# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the number of data points that are misclassified
error = sum(predictedLabels != 1)
# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)


```



















```{r}
library(e1071)
set.seed(7)
# prepare training scheme
control <- trainControl(method="repeatedcv", number=2, repeats=2)
# train the model
model <- train(Result ~ . - Season - Team1ID - Team2ID, data= trainingSet, method="lvq", preProcess="scale", trControl=control)
# estimate variable importance
importance <- varImp(model, scale=FALSE)
# summarize importance
print(importance)
# plot importance
```

```{r}
library(ggplot2)
plot(importance)
```


```{r}
head(trainingSet)
```