---
title: "Adrian March Madness"
output: html_notebook
---

### Model 1
### Season Data -> Predict -> March Madness Outcomes


```{r}
library(dplyr)
```


Load both season and tournament data

```{r}
marchResults <- read.csv("data/NCAATourneyDetailedResults.csv", header = TRUE)
head(marchResults)

seasonResults <- read.csv("data/RegularSeasonDetailedResults.csv", header = TRUE)
head(seasonResults)
```
Season - this is the year of the associated entry in Seasons.csv (the year in which the final tournament occurs)

DayNum - this integer always ranges from 0 to 132, and tells you what day the game was played on. It represents an offset from the "DayZero" date in the "Seasons.csv" file. For example, the first game in the file was DayNum=20. Combined with the fact from the "Seasons.csv" file that day zero was 10/29/1984 that year, this means the first game was played 20 days later, or 11/18/1984. There are no teams that ever played more than one game on a given date, so you can use this fact if you need a unique key (combining Season and DayNum and WTeamID). In order to accomplish this uniqueness, we had to adjust one game's date. In March 2008, the SEC postseason tournament had to reschedule one game (Georgia-Kentucky) to a subsequent day, so Georgia had to actually play two games on the same day. In order to enforce this uniqueness, we moved the game date for the Georgia-Kentucky game back to its original scheduled date.

WTeamID - this identifies the id number of the team that won the game, as listed in the "Teams.csv" file. No matter whether the game was won by the home team or visiting team, or if it was a neutral-site game, the "WTeamID" always identifies the winning team.
WScore - this identifies the number of points scored by the winning team.

LTeamID - this identifies the id number of the team that lost the game.

LScore - this identifies the number of points scored by the losing team. Thus you can be confident that WScore will be greater than LScore for all games listed.

NumOT - this indicates the number of overtime periods in the game, an integer 0 or higher.

WLoc - this identifies the "location" of the winning team. If the winning team was the home team, this value will be "H". If the winning team was the visiting team, this value will be "A". If it was played on a neutral court, then this value will be "N". Sometimes it is unclear whether the site should be considered neutral, since it is near one team's home court, or even on their court during a tournament, but for this determination we have simply used the Kenneth Massey data in its current state, where the "@" sign is either listed with the winning team, the losing team, or neither team. If you would like to investigate this factor more closely, we invite you to explore Data Section 3, which provides the city that each game was played in, irrespective of whether it was considered to be a neutral site.

WFGM - field goals made (by the winning team)
WFGA - field goals attempted (by the winning team)
WFGM3 - three pointers made (by the winning team)
WFGA3 - three pointers attempted (by the winning team)
WFTM - free throws made (by the winning team)
WFTA - free throws attempted (by the winning team)
WOR - offensive rebounds (pulled by the winning team)
WDR - defensive rebounds (pulled by the winning team)
WAst - assists (by the winning team)
WTO - turnovers committed (by the winning team)
WStl - steals (accomplished by the winning team)
WBlk - blocks (accomplished by the winning team)
WPF - personal fouls committed (by the winning team)



Create training data from season results (Only 2014 as of now)

```{r}
#subset 2014 data into new df
seasonResultsClean <- seasonResults[seasonResults$Season == "2014",]

#change the column names so that they are neutral (not indicative of winner or loser)
names(seasonResultsClean) <- gsub(x = names(seasonResultsClean), pattern = "WTeam", replacement = "Team1")
names(seasonResultsClean) <- gsub(x = names(seasonResultsClean), pattern = "LTeam", replacement = "Team2")
names(seasonResultsClean) <- gsub(x = names(seasonResultsClean), pattern = "W", replacement = "Team1")
names(seasonResultsClean) <- gsub(x = names(seasonResultsClean), pattern = "L", replacement = "Team2")

### Adjust the location variable for team 1 and create a location variable for team 2

#fix name
colnames(seasonResultsClean)[7] <- "Team1Loc"

#convert to character type for ease of processing
seasonResultsClean$Team1Loc <- as.character(seasonResultsClean$Team1Loc)

#create a location variable for team 2 that is the opposite of Team1Loc
seasonResultsClean <- seasonResultsClean %>% 
  mutate(Team2Loc = case_when(Team1Loc == "N"  ~ "N", 
                            Team1Loc == "H" ~ "A",
                            Team1Loc == "A" ~ "H", 
                            TRUE ~ as.character(Team1Loc)))

#factorize both columns again
seasonResultsClean$Team1Loc <- as.factor(seasonResultsClean$Team1Loc)
seasonResultsClean$Team2Loc <- as.factor(seasonResultsClean$Team2Loc)

#create a results variable
seasonResultsClean <- seasonResultsClean %>% 
  mutate(Result = 1)

#preview the data
head(seasonResultsClean)

### Duplicate and reverse half of the data

#duplicate data
seasonResultsCleanReverse <- seasonResultsClean

#reverse the result
seasonResultsCleanReverse <- seasonResultsCleanReverse %>%
     mutate(Result = replace(Result, Result == 1, 0)) %>%
     as.data.frame()

#rename Team 1 to Team 2 and vice versa
names(seasonResultsCleanReverse) <- gsub(x = names(seasonResultsCleanReverse), pattern = "1", replacement = "temp1")
names(seasonResultsCleanReverse) <- gsub(x = names(seasonResultsCleanReverse), pattern = "2", replacement = "temp2")
names(seasonResultsCleanReverse) <- gsub(x = names(seasonResultsCleanReverse), pattern = "temp1", replacement = "2")
names(seasonResultsCleanReverse) <- gsub(x = names(seasonResultsCleanReverse), pattern = "temp2", replacement = "1")

#reorder columns manually (TODO: this could probably be done better)
seasonResultsCleanReverse <- seasonResultsCleanReverse[,c(1,2,5,6,3,4,35,8,22,23,24,25,26,27,28,29,30,31,32,33,34,9,10,11,12,13,14,15,16,17,18,19,20,21,7,36)]

#bind datasets together
seasonResultsFinal <- bind_rows(seasonResultsClean,seasonResultsCleanReverse,id = NULL)

#drop certain columns in the results final data created above (TODO: these drops were selected based on current goals for the model and can be changed)
seasonResultsFinal <- seasonResultsFinal[,-c(2,7,8,35),drop=FALSE]
head(seasonResultsFinal)
```

```{r}
#create a list of all teams
teamIDList <- unique(seasonResultsFinal$Team1ID, incomparables = FALSE)

#create a vector of all columns that wish to be averaged
averageMeTeam1 <- c(3,6:18)
averageMeTeam2 <- c(5:19:31)

#average all of team 1 stats

#loop through all columns in the dataset
for(i in 1:ncol(seasonResultsFinal)){
  
  #loop through all teams
  for(j in 1:NROW(teamIDList)){
    
      #replace column value with column average
      if(i %in% averageMeTeam1 == TRUE){
      seasonResultsFinal[seasonResultsFinal$Team1ID == teamIDList[[j]],i] <- mean(seasonResultsFinal[seasonResultsFinal$Team1ID == teamIDList[[j]],i], na.rm = TRUE)
    }
  }
}

#loop through all columns in the dataset
for(i in 1:ncol(seasonResultsFinal)){
  
  #loop through all teams
  for(j in 1:NROW(teamIDList)){
    
      #replace column value with column average
      if(i %in% averageMeTeam2 == TRUE){
      seasonResultsFinal[seasonResultsFinal$Team2ID == teamIDList[[j]],i] <- mean(seasonResultsFinal[seasonResultsFinal$Team2ID == teamIDList[[j]],i], na.rm = TRUE)
    }
  }
}

head(seasonResultsFinal)

write.csv(seasonResultsFinal, file = "trainingdatav1.csv", row.names = FALSE)
```



Create testing data from tourney results
```{r}
#create a subset of only the 2014 results
tourneyResults2014 <- tourneyResults[tourneyResults$Season == "2014", c(1,3,4,5)]

#wrangle the data
# columns to paste together
cols <- c( "Season" , "WTeamID" , "LTeamID" )

# create a new column `ID` with the three columns collapsed together
tourneyResults2014$ID <- apply( tourneyResults2014[ , colNames ] , 1 , paste , collapse = "_" )

# remove the unnecessary columns
tourneyResults2014 <- tourneyResults2014[ , !( names( tourneyResults2014 ) %in% cols ) ]

#fill 
for (i in 1:NROW(tourneyResults2014)){
  tourneyResults2014[i,1] <- 1
}

#rename column
colnames(tourneyResults2014)[1] <- "Actual"

#Reorder columns
tourneyResults2014 <- tourneyResults2014[,c(2,1)]
```


