---
title: "Total Possessions"
author: "Lotan Weininger"
date: "March 5, 2018"
output: html_document
---
#Set Year = 2015
#Trying to identify total possessions/team in order to figure out possession differentials in matchups
#Read in Datasets
```{r}
#Events <- read.csv("data/Events_2015.csv", header = TRUE)
CompactResults <- read.csv("data/RegularSeasonCompactResults.csv", header = TRUE)
DetailedResults <- read.csv("data/RegularSeasonDetailedResults.csv", header = TRUE)

#head(Events)
head(CompactResults)
head(DetailedResults)
```


#Only want to deal with Detailed Results and actually compile all the data for an individual team for 2015
```{r}
DetailedResults2015 <- DetailedResults[DetailedResults$Season == "2015",]
head(DetailedResults)
```

Find all unique teams in the given season

```{r}
#find list of all teams during that season
teams <- data.frame(DetailedResults2015[unique(DetailedResults2015$WTeamID, incomparables = FALSE), 3])

#preview data
head(teams)

#count number of teams
nrow(teams)
```

#Read in Coach Data
```{r}
CoachData <- read.csv("data/TeamCoaches.csv", header = TRUE)
head(CoachData)
CoachRandomTeam <- CoachData[which(CoachData$Season == "1985" & CoachData$TeamID == "1106" & CoachData$LastDayNum == "154"), 5]
CoachRandomTeam
```

```{r}
#detailedresultsWin <- DetailedResults2015[DetailedResults2015$WTeamID == 1234,]
#detailedresultsLOSS <- DetailedResults2015[DetailedResults2015$LTeamID == 1234,]
#detailedresultsHomeW <- detailedresultsH[detailedresultsH$WLoc == "H",]
#detailedresultsHomeL <- detailedresultsLOSS[detailedresultsLOSS$WLoc == "A",]
#HomeW <- nrow(detailedresultsHomeW)
#HomeL <- nrow(detailedresultsHomeL)
#WinPerathome <- HomeW/(HomeW+HomeL)
#WinPerathome


#detailedresultsAwayW <- detailedresultsH[detailedresultsH$WLoc == "A",]
#detailedresultsAwayL <- detailedresultsLOSS[detailedresultsLOSS$WLoc == "H",]
#AwayW <- nrow(detailedresultsAwayW)
#AwayL <- nrow(detailedresultsAwayL)
#TotalAway <- AwayW + AwayL
#WinPeratAway <- AwayW/TotalAway
#WinPeratAway


#detailedresultsNeutW <- detailedresultsH[detailedresultsH$WLoc == "N",]
#detailedresultsNeutL <- detailedresultsLOSS[detailedresultsLOSS$WLoc == "N",]
#NeutW <- nrow(detailedresultsNeutW)
#NeutL <- nrow(detailedresultsNeutL)
#TotalNeut <- NeutW +NeutL
#WinPeratNeut <- NeutW/TotalNeut
#WinPeratNeut


#WinPerNotHome <- (NeutW + AwayW) / (TotalAway + TotalNeut)
#WinPerNotHome
```

```{r}



  for (j in 3:17){
    
    YearID <- paste("20")
    DetailedResults2015 <- DetailedResults[DetailedResults$Season == "2015",]
    head(DetailedResults)
    #find list of all teams during that season
    teams <- data.frame(DetailedResults2015[unique(DetailedResults2015$WTeamID, incomparables = FALSE), 3])

    #preview data
    head(teams)

    #count number of teams
    nrow(teams)
    for (i in 1:numTeams){
    numTeams <- nrow(teams)
    
  detailedResultsW <- DetailedResults2015[DetailedResults2015$WTeamID == teams[i,], c(1,2,3,4,8,9,10,11,12,13,14,15,16,17,18,19,20,21) ]
  
  detailedResultsL <- DetailedResults2015[DetailedResults2015$LTeamID == teams[i,], c(1,2,5,6,8,22,23,24,25,26,27,28,29,30,31,32,33,34) ]
  
  colnames(detailedResultsW) <- c("Season", "DayNum", "TeamID", "Score", "OT", "FGM", "FGA","FGM3", "FGA3", "FTM", "FTA", "OR", "DR", "AST", "TO", "STL", "BLK", "PF")
  colnames(detailedResultsL) <- c("Season", "DayNum", "TeamID", "Score", "OT", "FGM", "FGA","FGM3", "FGA3", "FTM", "FTA", "OR", "DR", "AST", "TO", "STL", "BLK", "PF")
  
  detailedBind <- rbind(detailedResultsW, detailedResultsL)
  detailedBind$Wins <- nrow(detailedResultsW)
  detailedBind$Losses <- nrow(detailedResultsL)
  detailedBind$STDV_Score <- sd(detailedBind$Score, na.rm = FALSE)
  detailedBind$STDV_FGM <- sd(detailedBind$FGM, na.rm = FALSE)
  detailedBind$STDV_FGA <- sd(detailedBind$FGA, na.rm = FALSE)
  detailedBind$STDV_FGM3 <- sd(detailedBind$FGM3, na.rm = FALSE)
  detailedBind$STDV_FGA3 <- sd(detailedBind$FGA3, na.rm = FALSE)
  detailedBind$STDV_FTM <- sd(detailedBind$FTM, na.rm = FALSE)
  detailedBind$STDV_FTA <- sd(detailedBind$FTA, na.rm = FALSE)
  detailedBind$STDV_OR <- sd(detailedBind$OR, na.rm = FALSE)
  detailedBind$STDV_DR <- sd(detailedBind$DR, na.rm = FALSE)
  detailedBind$STDV_AST <- sd(detailedBind$AST, na.rm = FALSE)
  detailedBind$STDV_TO <- sd(detailedBind$TO, na.rm = FALSE)
  detailedBind$STDV_STL <- sd(detailedBind$STL, na.rm = FALSE)
  detailedBind$STDV_BLK <- sd(detailedBind$BLK, na.rm = FALSE)
  detailedBind$STDV_PF <- sd(detailedBind$PF, na.rm = FALSE)

  detailedresultsWin <- DetailedResults2015[DetailedResults2015$WTeamID == teams[i,],]
  detailedresultsLOSS <- DetailedResults2015[DetailedResults2015$LTeamID == teams[i,],]
  detailedBind$Wins <- nrow(detailedresultsWin)
  detailedBind$Loss <- nrow(detailedresultsLOSS)
  detailedresultsHomeW <- detailedresultsWin[detailedresultsWin$WLoc == "H",]
  detailedresultsHomeL <- detailedresultsLOSS[detailedresultsLOSS$WLoc == "A",]
  HomeW <- nrow(detailedresultsHomeW)
  HomeL <- nrow(detailedresultsHomeL)
  detailedBind$WinPerathome <- HomeW/(HomeW+HomeL)
  detailedBind$WinPerathome


  detailedresultsAwayW <- detailedresultsWin[detailedresultsWin$WLoc == "A",]
  detailedresultsAwayL <- detailedresultsLOSS[detailedresultsLOSS$WLoc == "H",]
  AwayW <- nrow(detailedresultsAwayW)
  AwayL <- nrow(detailedresultsAwayL)
  TotalAway <- AwayW + AwayL
  detailedBind$WinPeratAway <- AwayW/TotalAway
  detailedBind$WinPeratAway


  detailedresultsNeutW <- detailedresultsWin[detailedresultsWin$WLoc == "N",]
  detailedresultsNeutL <- detailedresultsLOSS[detailedresultsLOSS$WLoc == "N",]
  NeutW <- nrow(detailedresultsNeutW)
  NeutL <- nrow(detailedresultsNeutL)
  TotalNeut <- NeutW +NeutL
  detailedBind$WinPeratNeut <- NeutW/TotalNeut
  detailedBind$WinPeratNeut


  detailedBind$WinPerNotHome <- (NeutW + AwayW) / (TotalAway + TotalNeut)
  detailedBind$WinPerNotHome
  

  #Coach <- CoachData[which(CoachData$Season == detailedBind$Season & CoachData$TeamID == detailedBind$TeamID & CoachData$LastDayNum == 154), 5]
   
  SeasonAverageTemp <- colMeans(detailedBind[c(1,3,4,6:38)])
  SeasonAverageTemp <- SeasonAverageTemp[c(1,2,17:18)]
  
  if (i == 1){
    seasonAverage <- as.data.frame(t(SeasonAverageTemp))
  } else{
    seasonAverage <- rbind(seasonAverage, as.data.frame(t(SeasonAverageTemp)) )
  }
  }
}
#print(Coach)
#print(detailedBind$Coach)
print(seasonAverage)
#print(detailedBind$)

write.csv(seasonAverage, file = "2015-teams-season-average.csv", row.names = FALSE)

```


```{r}
#DetailedResults1102W <- DetailedResults2015[DetailedResults2015$WTeamID == "1102", c(1,2,3,4,8,9,10,11,12,13,14,15,16,17,18,19,20,21) ]
#DetailedResults1102W
#DetailedResults1102L <- DetailedResults2015[DetailedResults2015$LTeamID == "1102", c(1,2,5,6,8,22,23,24,25,26,27,28,29,30,31,32,33,34) ]
#DetailedResults1102L
#colnames(DetailedResults1102W) <- c("Season", "DayNum", "TeamID", "Score", "OT", "FGM", "FGA","FGM3", "FGA3", "FTM", "FTA", "OR", "DR", "AST", "TO", "STL", "BLK", "PF")
#colnames(DetailedResults1102L) <- c("Season", "DayNum", "TeamID", "Score", "OT", "FGM", "FGA","FGM3", "FGA3", "FTM", "FTA", "OR", "DR", "AST", "TO", "STL", "BLK", "PF")
#DetailedResults1102L
#DetailedResults1102W
#Detailed1102 <- rbind(DetailedResults1102L, DetailedResults1102W)
#Detailed1102
```

#Get the averages of every column
```{r}
#SeasonAverage <- colMeans(Detailed1102[c(1,3,4,6,7,8,9,10,11,12,13,14,15,16,17,18)])
#SeasonAverage <- as.data.frame(t(SeasonAverage))
#SeasonAverage
```




###### Team Wins


```{r}
#Load libraries
library(dplyr)
#Load in the data 
seasonResults <- read.csv("data/RegularSeasonDetailedResults.csv", header = TRUE)
head(seasonResults)

seasonResultsOrignal1 <- seasonResults

```

```{r}
#change the column names so that they are identical for merging
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "WTeam", replacement = "Team")
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "LTeam", replacement = "Team")
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "W", replacement = "")
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "L", replacement = "")
```


```{r}
#combine WTeam and LTeam stats
seasonResultsCompiledWins <- bind_rows(seasonResults[,c(1,3)],seasonResults[,c(1,5)])
head(seasonResultsCompiledWins)
#find all unique values by BOTH Season and TeamID
#this will be the skeleton of the final database we want
seasonResultsCompiledUniqueWins <- distinct(seasonResultsCompiledWins,Season, TeamID, .keep_all = TRUE)
```

```{r}
#create new columns with names
seasonResultsCompiledUniqueWins$Wins <- NA
seasonResultsCompiledUniqueWins$Loses <- NA
seasonResultsCompiledUniqueWins$WinsAtHome <- NA
seasonResultsCompiledUniqueWins$WinsAtAway <- NA
seasonResultsCompiledUniqueWins$WinsAtNeutral <- NA
seasonResultsCompiledUniqueWins$LosesAtHome <- NA
seasonResultsCompiledUniqueWins$LosesAtAway <- NA
seasonResultsCompiledUniqueWins$LosesAtNeutral <- NA
```


```{r}
#loop through all the unique values
for (i in 1:NROW(seasonResultsCompiledUniqueWins)){
    
    #we want to replace the per game data in each row of the unique data frame with averages for that team in that season
    #first filter the FULL season results by two variables, the current row SEASON and TEAM ID
    #then only select the rows where an average can be computed
    #compute the average of the filtered and selected 14 columns and insert those 14 rows into the respective row of the unique data frame
    
    seasonResultsCompiledUniqueWins[i,3] <- nrow(seasonResultsOrignal1 %>%
                                               filter(Season == seasonResultsCompiledUniqueWins[i,1], WTeamID == seasonResultsCompiledUniqueWins[i,2]))
    
    seasonResultsCompiledUniqueWins[i,4] <- nrow(seasonResultsOrignal1 %>%
                                              filter(Season == seasonResultsCompiledUniqueWins[i,1], LTeamID == seasonResultsCompiledUniqueWins[i,2]))
    
    #HOME
    seasonResultsCompiledUniqueWins[i,5] <- nrow(seasonResultsOrignal1 %>%
                                               filter(Season == seasonResultsCompiledUniqueWins[i,1], WTeamID == seasonResultsCompiledUniqueWins[i,2], WLoc == "H"))
    
    #Away
    seasonResultsCompiledUniqueWins[i,6] <- nrow(seasonResultsOrignal1 %>%
                                              filter(Season == seasonResultsCompiledUniqueWins[i,1], WTeamID == seasonResultsCompiledUniqueWins[i,2], WLoc == "A"))
    
    #Neutral
    seasonResultsCompiledUniqueWins[i,7] <- nrow(seasonResultsOrignal1 %>%
                                               filter(Season == seasonResultsCompiledUniqueWins[i,1], WTeamID == seasonResultsCompiledUniqueWins[i,2], WLoc == "N"))
    
     #HOME
    seasonResultsCompiledUniqueWins[i,8] <- nrow(seasonResultsOrignal1 %>%
                                               filter(Season == seasonResultsCompiledUniqueWins[i,1], LTeamID == seasonResultsCompiledUniqueWins[i,2], WLoc == "H"))
    
    #Away
    seasonResultsCompiledUniqueWins[i,9] <- nrow(seasonResultsOrignal1 %>%
                                              filter(Season == seasonResultsCompiledUniqueWins[i,1], LTeamID == seasonResultsCompiledUniqueWins[i,2], WLoc == "A"))
    
    #Neutral
    seasonResultsCompiledUniqueWins[i,10] <- nrow(seasonResultsOrignal1 %>%
                                               filter(Season == seasonResultsCompiledUniqueWins[i,1], LTeamID == seasonResultsCompiledUniqueWins[i,2], WLoc == "N"))
}
```



#MUTATE FOR NEW COLUMNS --- Win%H Win%A Win%N Win%NotHome
```{r}
#seasonResultsCompiledUnique$WinPer <- NA

seasonResultsCompiledUniqueWins <- seasonResultsCompiledUniqueWins %>%
  mutate(WinPer = Wins / (Wins + Loses))
seasonResultsCompiledUniqueWins <- seasonResultsCompiledUniqueWins %>%
  mutate(WinPerHome = WinsAtHome / (WinsAtHome + LosesAtHome))
seasonResultsCompiledUniqueWins <- seasonResultsCompiledUniqueWins %>%
  mutate(WinPerAway = WinsAtAway / (WinsAtAway + LosesAtAway))
seasonResultsCompiledUniqueWins <- seasonResultsCompiledUniqueWins %>%
  mutate(WinPerNeut = WinsAtNeutral / (WinsAtNeutral + LosesAtNeutral))
seasonResultsCompiledUniqueWins <- seasonResultsCompiledUniqueWins %>%
  mutate(WinPerNotHome = (WinsAtNeutral + WinsAtAway) / ((WinsAtNeutral + LosesAtNeutral) + (WinsAtAway + LosesAtAway)))


head(seasonResultsCompiledUniqueWins)
```

```{r}
#nrow(seasonResultsOrignal %>%
 #        filter(Season == 2013, WTeamID == 1104))
```


```{r}
#write to file
write.csv(seasonResultsCompiledUniqueWins, file = "allteams-allseason-winloss.csv", row.names = FALSE)
```

#### SD
```{r}
#Load libraries
library(dplyr)
#Load in the data 
seasonResults <- read.csv("data/RegularSeasonDetailedResults.csv", header = TRUE)
head(seasonResults)

```

```{r}
#change the column names so that they are identical for merging
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "WTeam", replacement = "Team")
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "LTeam", replacement = "Team")
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "W", replacement = "")
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "L", replacement = "")
```

```{r}
#combine WTeam and LTeam stats
seasonResultsCompiled <- bind_rows(seasonResults[,c(1,3:4,9:21)],seasonResults[,c(1,5:6,22:34)])
head(seasonResultsCompiled)
#find all unique values by BOTH Season and TeamID
#this will be the skeleton of the final database we want
seasonResultsCompiledUnique <- distinct(seasonResultsCompiled,Season, TeamID, .keep_all = TRUE)
```

```{r}
#loop through all the unique values
library(matrixStats)
for (i in 1:NROW(seasonResultsCompiledUnique)){
    
    #we want to replace the per game data in each row of the unique data frame with averages for that team in that season
    #first filter the FULL season results by two variables, the current row SEASON and TEAM ID
    #then only select the rows where an average can be computed
    #compute the average of the filtered and selected 14 columns and insert those 14 rows into the respective row of the unique data frame
    seasonResultsCompiledUnique[i,c(3:16)] <- sapply((seasonResultsCompiled %>%
      filter(Season == seasonResultsCompiledUnique[i,1], TeamID == seasonResultsCompiledUnique[i,2]) %>%
      select(-(Season:TeamID))), sd)
}
```

```{r}
#write to file
write.csv(seasonResultsCompiledUnique, file = "allteams-allseason-SD.csv", row.names = FALSE)
```


#Possession Time

```{r}
#Agrregate all 7 CSV into one
Events2010 <- read.csv("data/Events_2010.csv", header = TRUE)
Events2011 <- read.csv("data/Events_2011.csv", header = TRUE)
Events2012 <- read.csv("data/Events_2012.csv", header = TRUE)
Events2013 <- read.csv("data/Events_2013.csv", header = TRUE)
Events2014 <- read.csv("data/Events_2014.csv", header = TRUE)
Events2015 <- read.csv("data/Events_2015.csv", header = TRUE)
Events2016 <- read.csv("data/Events_2016.csv", header = TRUE)
Events2017 <- read.csv("data/Events_2017.csv", header = TRUE)
Events2018 <- read.csv("data/Events_2018.csv", header = TRUE)
```


```{r}
#Load libraries
library(dplyr)
#Load in the data 
#seasonResults <- read.csv("../data/RegularSeasonDetailedResults.csv", header = TRUE)
#head(seasonResults)

```

```{r}
#change the column names so that they are identical for merging
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "WTeam", replacement = "Team")
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "LTeam", replacement = "Team")
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "W", replacement = "")
names(seasonResults) <- gsub(x = names(seasonResults), pattern = "L", replacement = "")
```

```{r}
#combine WTeam and LTeam stats
seasonResultsCompiled <- bind_rows(Events2010[,c(2,4:5,8:9)],Events2011[,c(2,4:5,8:9)],Events2012[,c(2,4:5,8:9)],Events2013[,c(2,4:5,8:9)],Events2014[,c(2,4:5,8:9)],Events2015[,c(2,4:5,8:9)],Events2016[,c(2,4:5,8:9)],Events2017[,c(2,4:5,8:9)],Events2018[,c(2,4:5,8:9)])
head(seasonResultsCompiled)
#find all unique values by BOTH Season and TeamID
#this will be the skeleton of the final database we want
seasonResultsCompiledUnique <- distinct(seasonResultsCompiled,Season, EventTeamID, .keep_all = TRUE)
```

```{r}
seasonResultsCompiled$TimeBetweenEvents <- ave(seasonResultsCompiled$ElapsedSeconds, FUN = function(x) c(0,diff(x)))
seasonResultsCompiled
```
```{r}
seasonResultsCompiled$TimeBetweenEvents[seasonResultsCompiled$TimeBetweenEvents < 0] <- 0
seasonResultsCompiled
```
```{r}
#Create new columns
seasonResultsCompiledUnique$TotalPossessionTime <- NA
```

```{r}

```

```{r}
#loop through all the unique values
#library(matrixStats)
for (i in 1:NROW(seasonResultsCompiledUnique)){
    
    #we want to replace the per game data in each row of the unique data frame with averages for that team in that season
    #first filter the FULL season results by two variables, the current row SEASON and TEAM ID
    #then only select the rows where an average can be computed
    #compute the average of the filtered and selected 14 columns and insert those 14 rows into the respective row of the unique data frame
    seasonResultsCompiledUnique[i,c(6)] <- sum(seasonResultsCompiled %>%
      filter(Season == seasonResultsCompiledUnique[i,1], EventTeamID == seasonResultsCompiledUnique[i,5]) %>%
      select(-(Season:EventTeamID)), na.rm = FALSE)
}
```

```{r}
WinData <- read.csv("data/RegularSeasonDetailedResults.csv", header = TRUE)
head(WinData)
```

```{r}
nrow(WinData %>%
       filter(Season == "2010", WTeamID == "1346"))
```
```{r}
seasonResultsCompiledUnique$Wins <- NA
seasonResultsCompiledUnique$Loses <- NA
```

```{r}
for (i in 1:NROW(seasonResultsCompiledUnique)){
    
    #we want to replace the per game data in each row of the unique data frame with averages for that team in that season
    #first filter the FULL season results by two variables, the current row SEASON and TEAM ID
    #then only select the rows where an average can be computed
    #compute the average of the filtered and selected 14 columns and insert those 14 rows into the respective row of the unique data frame
    seasonResultsCompiledUnique[i,c(7)] <- NROW(WinData %>%
      filter(Season == seasonResultsCompiledUnique[i,1], WTeamID == seasonResultsCompiledUnique[i,5]))
    
     seasonResultsCompiledUnique[i,c(8)] <- NROW(WinData %>%
      filter(Season == seasonResultsCompiledUnique[i,1], LTeamID == seasonResultsCompiledUnique[i,5]))
}
```

```{r}
seasonResultsCompiledUnique <- seasonResultsCompiledUnique %>%
  mutate(PossessionTimePerGame = TotalPossessionTime / (Wins + Loses))
```

```{r}
seasonResultsCompiledUnique <- seasonResultsCompiledUnique %>%
  mutate(PossessionTimePerGameinMinutes = PossessionTimePerGame / 60)
```

```{r}
seasonResultsCompiledUnique$WTeamID <- NULL
seasonResultsCompiledUnique$LTeamID <- NULL
seasonResultsCompiledUnique$ElapsedSeconds <- NULL
seasonResultsCompiledUnique$Wins <- NULL
seasonResultsCompiledUnique$Loses <- NULL
```

```{r}
names(seasonResultsCompiledUnique) <- c("Season", "TeamID", "TotalPossessionTime", "PossessionTimePerGame(S)","PossessionTimePerGame(M)")
```

```{r}
#write to file
write.csv(seasonResultsCompiledUnique, file = "allteams-allseason-AveragePossessionsperGame.csv", row.names = FALSE)
```
```

