---
title: "Jeet - Experimenting"
output: html_notebook
---


Set testing year
```{r}
library(stringr)
year<-2014
```


Pull seed data and tourney results
```{r}
# load data need data files Seeds, Results, Conferences, and Coaches
seeds <- read.csv("data/NCAATourneySeeds.csv")
results <- read.csv("data/RegularSeasonCompactResults.csv")
results<-subset(results,Season<year)
testData<-read.csv("data/SampleSubmissionStage1.csv")
```

```{r}
head(seeds)
head(results)

# Extracts seed number
seeds$Seed <- as.integer(str_extract_all(seeds$Seed, "[0-9]+"))
```

```{r}
#Taking out results before 2014 for Round 1 Prediction (2014-2017)
results <- subset(results,Season < 2014)
head(results)

library(dplyr)

# Copies the results into 
transformed1 <- results
transformed2 <- results

# Changes format to be similar to test file
transformed1 <- mutate(transformed1,
  team1_id = WTeamID,
  team2_id = LTeamID,
  result = 1)

transformed2 <- mutate(transformed2,
  team1_id = LTeamID,
  team2_id = WTeamID,
  result = 0)

# Gets only relevant data
transformed1 <- transformed1[c("Season","team1_id", "team2_id", "result")]
transformed2 <- transformed2[c("Season","team1_id", "team2_id", "result")]

# Doubles the rows to switch around team1 and team 2 and include a loss
training <- union_all(transformed1,transformed2)
```


```{r}
# Copies seeds into other databases for merging
seeds1 <- seeds
names(seeds1) <- c("Season", "team1_Seed", "team1_id")
seeds2 <- seeds
names(seeds2) <- c("Season", "team2_Seed", "team2_id")

# Left outer joins to associate seed data
training <- merge(x = training, y = seeds1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = seeds2, by = c("Season", "team2_id"), all.x = TRUE)
```


```{r}
# Locates all na's in data and replaces them with 20's
training$team1_Seed <- ifelse(is.na(training$team1_Seed), 20, training$team1_Seed)
training$team2_Seed <- ifelse(is.na(training$team2_Seed), 20, training$team2_Seed)

summary(training)
```

Load Coach names and format accordingly
```{r}
coaches <- read.csv("data/TeamCoaches.csv")
coaches <- subset(coaches, LastDayNum==154)
coaches <- coaches[c("Season", "TeamID", "CoachName")]
coaches$CoachName <- as.character(coaches$CoachName)
coaches1 <- coaches
names(coaches1) <- c("Season", "team1_id", "team1_Coach")
coaches2 <- coaches
names(coaches2) <- c("Season", "team2_id", "team2_Coach")
```


```{r}
# Left outer joins to associate coach data
training <- merge(x = training, y = coaches1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = coaches2, by = c("Season", "team2_id"), all.x = TRUE)
```


Load Conference Data and format accordingly
```{r}
conference <- read.csv("data/TeamConferences.csv")
conference$ConfAbbrev <- as.character(conference$ConfAbbrev)
conference1 <- conference
names(conference1) <- c("Season", "team1_id", "team1_conference")
conference2 <- conference
names(conference2) <- c("Season", "team2_id", "team2_conference")
```

```{r}
<<<<<<< HEAD
rankings1 <- rankings
names(rankings1) <- c("Season", "team1_id", "team1_avgranking", "team1_sdranking")
rankings2 <- rankings
names(rankings2) <- c("Season", "team2_id", "team2_avg_ranking", "team2_sd_ranking")

# Left outer joins to associate ranking data
training <- merge(x = training, y = rankings1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = rankings2, by = c("Season", "team2_id"), all.x = TRUE)
rm(rankings, rankings1, rankings2)
```


Merging in Win/Loss Record
```{r}
winloss <- read.csv("compiled-features/allteams-allseason-winloss.csv")
winloss <- winloss[,c(1:2, 11:15)]

winloss1 <- winloss
names(winloss1)[2] <- "team1_id"
names(winloss1)[3:7] <- paste(sep = "_", "team1", names(winloss1)[3:7])

winloss2 <- winloss
names(winloss2)[2] <- "team2_id"
names(winloss2)[3:7] <- paste(sep = "_", "team2", names(winloss1)[3:7])

# Left outer joins to associate win/loss data
training <- merge(x = training, y = winloss1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = winloss2, by = c("Season", "team2_id"), all.x = TRUE)
rm(winloss, winloss1, winloss2)
```


Merging in Average Stats
```{r}
<<<<<<< HEAD
# Selecting the scope of the data
#data2003 <- subset(data, Season < 2014)
#training2 <- subset(data, Season == 2014)
#training2 <- subset(training2, post == 0)

#data2003 <- union_all(data2003, training2)
#rm(training2)

data <- na.omit(data)
=======
avgStats <- read.csv("compiled-features/allteams-allseason-averages.csv")

avgStats1 <- avgStats
names(avgStats1)[2] <- "team1_id"
names(avgStats1)[3:16] <- paste(sep = "_", "team1", names(avgStats1)[3:16])

avgStats2 <- avgStats
names(avgStats2)[2] <- "team2_id"
names(avgStats2)[3:16] <- paste(sep = "_", "team2", names(avgStats2)[3:16])

# Left outer joins to associate win/loss data
training <- merge(x = training, y = avgStats1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = avgStats2, by = c("Season", "team2_id"), all.x = TRUE)
rm(avgStats, avgStats1, avgStats2)
```

Merging in Standard Deviation of Stats
```{r}
sdStats <- read.csv("compiled-features/allteams-allseason-SD.csv")

sdStats1 <- sdStats
names(sdStats1)[2] <- "team1_id"
names(sdStats1)[3:16] <- paste(sep = "_", "team1_sd_", names(sdStats1)[3:16])

sdStats2 <- sdStats
names(sdStats2)[2] <- "team2_id"
names(sdStats2)[3:16] <- paste(sep = "_", "team2_sd_", names(sdStats2)[3:16])

# Left outer joins to associate win/loss data
training <- merge(x = training, y = sdStats1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = sdStats2, by = c("Season", "team2_id"), all.x = TRUE)
rm(sdStats, sdStats1, sdStats2)
```

Merging in Years of Experience
```{r}
Stats <- read.csv("compiled-features/experience.csv")
Stats[,1] <- NULL

Stats1 <- Stats
names(Stats1)<- c("Season", "team1_id", "team1_avgExp", "team1_sdExp")

Stats2 <- Stats
names(Stats2)<- c("Season", "team2_id", "team2_avgExp", "team2_sdExp")

# Left outer joins to associate win/loss data
training <- merge(x = training, y = Stats1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = Stats2, by = c("Season", "team2_id"), all.x = TRUE)
rm(Stats, Stats1, Stats2)
```


Merging in Pts Against data
```{r}
Stats <- read.csv("compiled-features/ptsAgainst.csv")
Stats[,1] <- NULL

Stats1 <- Stats
names(Stats1)<- c("Season", "team1_id", "team1_avg_pts_against", "team1_sd_pts_against")

Stats2 <- Stats
names(Stats2)<- c("Season", "team2_id", "team2_avg_pts_against", "team2_sd_pts_against")

# Left outer joins to associate win/loss data
training <- merge(x = training, y = Stats1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = Stats2, by = c("Season", "team2_id"), all.x = TRUE)
rm(Stats, Stats1, Stats2)
```

```{r}
write.csv(training, "trainingSet.csv")
>>>>>>> parent of ea034fa... Changed Summary File
```

This section of code takes all the columns from teams 1 and 2, adds them together and calculates how much team1's numbers represent as a proportion of the sum
```{r}
<<<<<<< HEAD
x <- names(data)

x1 <- unlist(subset(x, grepl("team1",x)))
x2 <- unlist(subset(x, grepl("team2",x)))

dataSet <- data[,2:6]

len <- length(x1)


for(i in 2:len){
  column <- substr(x1[i], 7, nchar(x1[i]))
  c1 <- x1[i]
  c2 <- x2[i]
  #Added a 0.0000001 in the denominator to prevent division errors
  dataSet[,column] <- data[c1]/(data[c1]+data[c2]+0.000001)
}

dataSet <- dataSet[dataSet$post==1,]
rm(x, x1, x2, c1, c2)
=======
# Left outer joins to associate coach data
training <- merge(x = training, y = conference1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = conference2, by = c("Season", "team2_id"), all.x = TRUE)
>>>>>>> parent of 78baa00... Added Merged CSV
```


Normalize the feature via Z-score
```{r}
col <- dim(dataSet)[2]
normData <- dataSet
for(i in 7:col){
  normData[,i] = (normData[,i] - mean(normData[,i]))/sd(normData[,i])
}
=======
training$team1_Seed = (training$team1_Seed - mean(training$team1_Seed))/sd(training$team1_Seed)

training$team2_Seed = (training$team2_Seed - mean(training$team2_Seed))/sd(training$team2_Seed)
>>>>>>> parent of ea034fa... Changed Summary File
```

Split the data into a training and a testing set
```{r}
numSamples <- dim(normData)[1]
sampling.rate = 0.8

trainingIndex <- sample(1:numSamples, numSamples*sampling.rate, replace = FALSE)
trainingSet <- normData[trainingIndex,]

testingIndex <- setdiff(1:numSamples, trainingIndex)
testingSet <- normData[testingIndex,]
```


Train NN
```{r}
library(neuralnet)
<<<<<<< HEAD
f <- paste(names(trainingSet)[6:col], collapse = '+')
f<- paste('result~', f)
#f <- as.formula('result~.-Season-team2_id-team1_id-post')
nn <- neuralnet(f, data=trainingSet, hidden=c(18,4),linear.output=FALSE)
=======
f <- as.formula('result~team1_Seed+team2_Seed+team1_Coach+team2_Coach+team1_conference+team2_conference')
nn <- neuralnet(f, data=training, hidden=c(5,3),linear.output=FALSE)
>>>>>>> parent of ea034fa... Changed Summary File
```

Predict results of Testing Set
```{r}
predictions <- compute(nn,testingSet[,6:43])

predictions <- predictions$net.result
roundedPredictions <- round(predictions)

sizeTest <- dim(predictions)[1]
error = sum(roundedPredictions != testingSet$result)
misclassification = error / sizeTest

misclassification
```


