---
title: "Jeet - Experimenting"
output: html_notebook
---


Set testing year
```{r}
library(stringr)
year<-2019
```


Pull seed data and tourney results
```{r}
# load data need data files Seeds, Results, Conferences, and Coaches
seeds <- read.csv("../data/NCAATourneySeeds.csv")
regResults <- read.csv("../data/RegularSeasonCompactResults.csv")
regResults <- regResults[c("Season", "WTeamID", "LTeamID")]
regResults$post <- 0

postResults <- read.csv("../data/NCAATourneyCompactResults.csv")
postResults <- postResults[c("Season", "WTeamID", "LTeamID")]
postResults$post <- 1


otherResults <- read.csv("../data/SecondaryTourneyCompactResults.csv")
otherResults <- otherResults[c("Season", "WTeamID", "LTeamID")]
otherResults$post <- 0

results <- union_all(regResults, postResults, otherResults)
results<-subset(results,Season<year)
testData<-read.csv("../data/SampleSubmissionStage1.csv")
rm(otherResults, postResults, regResults)
```

```{r}
head(seeds)
head(results)

# Extracts seed number
seeds$Seed <- as.integer(str_extract_all(seeds$Seed, "[0-9]+"))
```

```{r}
library(dplyr)

# Copies the results into 
transformed1 <- results
transformed2 <- results

# Changes format to be similar to test file
transformed1 <- mutate(transformed1,
  team1_id = WTeamID,
  team2_id = LTeamID,
  result = 1)

transformed2 <- mutate(transformed2,
  team1_id = LTeamID,
  team2_id = WTeamID,
  result = 0)

# Gets only relevant data
transformed1 <- transformed1[c("Season","post", "team1_id", "team2_id", "result")]
transformed2 <- transformed2[c("Season","post", "team1_id", "team2_id", "result")]

# Doubles the rows to switch around team1 and team 2 and include a loss
training <- union_all(transformed1,transformed2)
rm(transformed1, transformed2)
```


```{r}
# Copies seeds into other databases for merging
seeds1 <- seeds
names(seeds1) <- c("Season", "team1_Seed", "team1_id")
seeds2 <- seeds
names(seeds2) <- c("Season", "team2_Seed", "team2_id")

# Left outer joins to associate seed data
training <- merge(x = training, y = seeds1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = seeds2, by = c("Season", "team2_id"), all.x = TRUE)
rm(seeds, seeds1, seeds2)
```


```{r}
# Locates all na's in data and replaces them with 20's
training$team1_Seed <- ifelse(is.na(training$team1_Seed), 20, training$team1_Seed)
training$team2_Seed <- ifelse(is.na(training$team2_Seed), 20, training$team2_Seed)

summary(training)
```

Load Coach names and format accordingly
```{r}
coaches <- read.csv("allteams-allseason-allcoaches-winloss.csv")
coaches <- coaches[c("Season", "TeamID", "CoachYears", "CoachWinPer")]
coaches1 <- coaches
names(coaches1) <- c("Season", "team1_id", "team1_CoachYears", "team1_CoachWinPer")
coaches2 <- coaches
names(coaches2) <- c("Season", "team2_id","team2_CoachYears", "team2_CoachWinPer")
```


```{r}
# Left outer joins to associate coach data
training <- merge(x = training, y = coaches1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = coaches2, by = c("Season", "team2_id"), all.x = TRUE)
rm(coaches, coaches1, coaches2)
```


-- DELETED DUE TO LARGE NUMBER OF CONFERENCES --
Load Conference Data and dummy it

conference <- read.csv("data/TeamConferences.csv")
library(dummies)
conference <- cbind(conference, dummy(conference$ConfAbbrev))

conference1 <- conference
names(conference1) <- c("Season", "team1_id", "team1_conference")
conference2 <- conference
names(conference2) <- c("Season", "team2_id", "team2_conference")




# Left outer joins to associate conference data
training <- merge(x = training, y = conference1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = conference2, by = c("Season", "team2_id"), all.x = TRUE)
rm(conference, conference1, conference2)


Investigating the Massey Ordinals Data
```{r}
massey <- read.csv("../data/MasseyOrdinals.csv")
massey <- subset(massey, RankingDayNum == 133)

rankings <- aggregate(massey[,5],list(massey$Season, massey$TeamID),mean)
sd <- aggregate(massey[,5],list(massey$Season, massey$TeamID),sd)
pom <- subset(massey, SystemName == "POM")
pom$RankingDayNum = NULL
pom$SystemName = NULL
names(pom)[3] <- "pom"

names(rankings) <- c("Season", "TeamID", "mean")
rankings$sd <- sd$x

rankings <- merge(x = rankings, y = pom, by = c("Season", "TeamID"), all.x = TRUE)
# takes days without POM rankings and replaces with average ranking)
rankings$pom <- ifelse(is.na(rankings$pom), rankings$mean, rankings$pom)
rm(sd, massey)
```

```{r}
rankings1 <- rankings
names(rankings1) <- c("Season", "team1_id", "team1_avgranking", "team1_sdranking", "team1_pomranking")
rankings2 <- rankings
names(rankings2) <- c("Season", "team2_id", "team2_avg_ranking", "team2_sd_ranking", "team2_pomranking")

# Left outer joins to associate ranking data
training <- merge(x = training, y = rankings1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = rankings2, by = c("Season", "team2_id"), all.x = TRUE)
rm(rankings, rankings1, rankings2)
```


Merging in Win/Loss Record
```{r}
winloss <- read.csv("allteams-allseason-winloss.csv")
winloss <- winloss[,c(1:2, 11:15)]

winloss1 <- winloss
names(winloss1)[2] <- "team1_id"
names(winloss1)[3:7] <- paste(sep = "_", "team1", names(winloss1)[3:7])

winloss2 <- winloss
names(winloss2)[2] <- "team2_id"
names(winloss2)[3:7] <- paste(sep = "_", "team2", names(winloss2)[3:7])

# Left outer joins to associate win/loss data
training <- merge(x = training, y = winloss1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = winloss2, by = c("Season", "team2_id"), all.x = TRUE)
rm(winloss, winloss1, winloss2)
```


Merging in Average Stats
```{r}
avgStats <- read.csv("allteams-allseason-averages.csv")

avgStats1 <- avgStats
names(avgStats1)[2] <- "team1_id"
names(avgStats1)[3:16] <- paste(sep = "_", "team1", names(avgStats1)[3:16])

avgStats2 <- avgStats
names(avgStats2)[2] <- "team2_id"
names(avgStats2)[3:16] <- paste(sep = "_", "team2", names(avgStats2)[3:16])

# Left outer joins to associate win/loss data
training <- merge(x = training, y = avgStats1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = avgStats2, by = c("Season", "team2_id"), all.x = TRUE)
rm(avgStats, avgStats1, avgStats2)
```

Merge in Sum Stats
```{r}
sumStats <- read.csv("allteams-allseason-sum.csv")
sumStats<-sumStats[,c(1:2,17,32:35)]

head(sumStats)


sumStats1 <- sumStats
names(sumStats1)[2] <- "team1_id"
names(sumStats1)[3:7] <- paste(sep = "_", "team1", names(sumStats1)[3:7])

sumStats2 <- sumStats
names(sumStats2)[2] <- "team2_id"
names(sumStats2)[3:7] <- paste(sep = "_", "team2", names(sumStats2)[3:7])

# Left outer joins to associate win/loss data
training <- merge(x = training, y = sumStats1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = sumStats2, by = c("Season", "team2_id"), all.x = TRUE)
rm(sumStats, sumStats1, sumStats2)

```


Merging in Standard Deviation of Stats
```{r}
sdStats <- read.csv("allteams-allseason-SD.csv")

sdStats1 <- sdStats
names(sdStats1)[2] <- "team1_id"
names(sdStats1)[3:16] <- paste(sep = "_", "team1_sd_", names(sdStats1)[3:16])

sdStats2 <- sdStats
names(sdStats2)[2] <- "team2_id"
names(sdStats2)[3:16] <- paste(sep = "_", "team2_sd_", names(sdStats2)[3:16])

# Left outer joins to associate win/loss data
training <- merge(x = training, y = sdStats1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = sdStats2, by = c("Season", "team2_id"), all.x = TRUE)
rm(sdStats, sdStats1, sdStats2)
```

Merging in Years of Experience
```{r}
Stats <- read.csv("experience.csv")
Stats[,1] <- NULL

Stats1 <- Stats
names(Stats1)<- c("Season", "team1_id", "team1_avgExp", "team1_sdExp")

Stats2 <- Stats
names(Stats2)<- c("Season", "team2_id", "team2_avgExp", "team2_sdExp")

# Left outer joins to associate win/loss data
training <- merge(x = training, y = Stats1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = Stats2, by = c("Season", "team2_id"), all.x = TRUE)
rm(Stats, Stats1, Stats2)
```


Merging in Pts Against data
```{r}
Stats <- read.csv("ptsAgainst.csv")
Stats[,1] <- NULL

Stats1 <- Stats
names(Stats1)<- c("Season", "team1_id", "team1_avg_pts_against", "team1_sd_pts_against")

Stats2 <- Stats
names(Stats2)<- c("Season", "team2_id", "team2_avg_pts_against", "team2_sd_pts_against")

# Left outer joins to associate win/loss data
training <- merge(x = training, y = Stats1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = Stats2, by = c("Season", "team2_id"), all.x = TRUE)
rm(Stats, Stats1, Stats2)
```

Merging in Pts per Possession
```{r}
Stats <- read.csv("allteams-allseason-AveragePossessionsperGame.csv")
Stats <- Stats[,c(1,2,5)]

Stats1 <- Stats
names(Stats1)<- c("Season", "team1_id", "team1_avg_pospergame")

Stats2 <- Stats
names(Stats2)<- c("Season", "team2_id",  "team2_avg_pospergame")

# Left outer joins to associate win/loss data
training <- merge(x = training, y = Stats1, by = c("Season", "team1_id"), all.x = TRUE)
training <- merge(x = training, y = Stats2, by = c("Season", "team2_id"), all.x = TRUE)
rm(Stats, Stats1, Stats2)
```


```{r}
training
```

Create EFGper, TSper, REBper column
```{r}
#Create EFGper column
#Note: eFG % = [FGM + 0.5 * FGM3] / FGA
training$team1_EFGper<-((training$team1_FGM+0.5*training$team1_FGM3)/(training$team1_FGA))
training$team2_EFGper<-((training$team2_FGM+0.5*training$team2_FGM3)/(training$team2_FGA))

#Create TSper column
#Note: TSper = [ Score / [2*[FGA+[0.44*FTA]]]
training$team1_TSper<-(training$team1_Score/(2*(training$team1_FGA+(0.44*training$team1_FTA))))
training$team2_TSper<-(training$team2_Score/(2*(training$team2_FGA+(0.44*training$team2_FTA))))


```


```{r}
fuck <- subset(training, Season == 2018)
```

Write to CSV
```{r}
write.csv(training, "trainingSet.csv")
```
